### 2014-11-12 contract-language.tex
hfill before qed

The syntax of our core contract language is shown in Figure
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed and untyped effect variables.
We use a special effect variable ($\cureff$) to denote the effect of
\emph{current operation} - the operation for which a contract is being written.
Notice that $\cureff$ occurs free in the contract. We will fix its scope when
classifying the contracts (\S~\ref{sec:classify}). The type of an effect is
simply the name of the operation (eg: \cf{withdraw}) that induced the effect.
We admit disjunction in types to let an effect variable range over multiple
operation names. The contract $\small \forall (a : \tau_1 \vee \tau_2).~\psi$
is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1} \vee
\oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect variable ranges over all
operation names.

@@

\subsection{Contract Classification}
\label{sec:classify}

@@

The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
$\ecc$ contracts in all well-formed executions, then the same store, extended
with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
all well-formed \name contracts. The proof of the theorem is given in the
supplementary material\footnote{The supplementary material also provides the
concrete reduction rules for enforcing the consistency classes.}.

>>>>

The syntax of our core contract language is shown in Figure
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed and untyped effect variables.
We use a special effect variable ($\cureff$) to denote the effect of
\emph{current operation} - the operation for which a contract is being written.
The type of an effect is simply the name of the operation (eg: \cf{withdraw})
that induced the effect. We admit disjunction in types to let an effect variable
range over multiple operation names. The contract $\small \forall (a : \tau_1
\vee \tau_2).~\psi$ is just syntactic sugar for $\small \forall a.
(\oper{a}{\tau_1} \vee \oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect
variable ranges over all operation names.

@@

\subsection{Contract Comparison and Classification}

@@

\noindent The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
store, extended with the classification scheme shown in
Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
proof of the theorem is given below:

\begin{proof}
  Hypothesis:
  \begin{smathpar}
  \begin{array}{cr}
    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E', \sigma \pll \Sigma} & H\npp\\
    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
  \end{array}
  \end{smathpar}
  Since $\tau$ is the contract class of $\cv$, by inversion, we have
  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
  \end{array}
  \end{smathpar}
   Since $\eff$ denotes new effect, it is a fresh variable that does
   not occur free in $\Delta$. From $H3$, after instantiating bound
   $\cureff$ with $\eff$, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
      & H\npp\\
  \end{array}
  \end{smathpar}
  Due to the soundness of natural deduction for first-order logic,
  $H4$ implies that for all models $\mathcal{M}$ such that
  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \E' \models \Delta & H\npp\\
  \end{array}
  \end{smathpar}
  Proof follows from $H1$, $H5$, and $H4$.
  \hfill \qed
\end{proof}

It is important to note that Theorem~\ref{thm:classification-sound}
does not ascribe any semantics to the reduction relation
($\xrightarrow{}$). As such, it makes no assumptions about how the
store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
guarantees. The specific implementation strategy is determined by the
operational semantics of the store, which \emph{defines} the reduction
relation for that particular store. The following section describes
operational semantics of the store used by the \name implementation.
